Invasive lionfish present region-wide variation of allometric growth in the Western Atlantic

### 2018-05-09 docs/Manuscript.tex
Updated numbers across manuscript and added more discussion

\begin{abstract}
Lionfish (\emph{Pterois volitans/miles}) are an invasive species in the
Western Atlantic and the Caribbean. In order to better manage the
invasion, we must be able to accurately estimate their total biomass.
This work compares length-weight relationships of the invasive lionfish
through the invasion range. A review of 15 length-weight relationships
reported in 11 peer-reviewed studies shows that lionfish exhibit spatial
variation in weight-at-length. The reviewed parameters indicate that,
for the same length, lionfish in the Caribbean have lower body mass than
in the Atlantic or Gulf of Mexico. This highlights the importance of
using site-specific parameters to estimate biomass from length
observations. This study also reports a new pair of length-weight
parameters (\(a = 3.2056 imes 10^{-6}; b = 3.235\)) for organisms
sampled in the central Mexican Caribbean region. These findings can have
major implications in management, especially when estimating biomass
available for harvest, predicting effects on local ecosystems, or
evaluating the effectiveness of removal programs.
\end{abstract}

@@

Empirical modeling efforts examining the feasibility of establishing
fisheries for lionfish involve modeling changes in biomass in response
to changes in mortality (\emph{i.e.} culling
\citep{cote_2014,barbour_2011,andradibrown_2017}). In this case, biomass
represents the sum of all individual fish weights. The individual weight
(\(W\)) of an organism can be estimated from its length using the
allometric growth equation, \(W = aL^b\), where \(L\) is an organism's
length and \(a\) and \(b\) are constants. This length-weight
relationship for any given species is an essential component of
biomass-based modeling, and can vary across regions as a response to
biotic (\emph{e.g.} local food availability) and abiotic (\emph{e.g.}
water temperature) conditions \citep{johnson_2016}. Site-specific
parameters are therefore necessary in order to obtain accurate biomass
estimates in a given region, especially when research involves
identifying the total biomass available for harvest by fishers
\citep{barbour_2011,morris_2011,johnston_2015,chin_2016}. Equation
parameters for lionfish exist for North Carolina, Northern and Southern
Gulf of Mexico, the Southern Mexican Caribbean, Bahamas, Little Cayman,
Jamaica, Bonaire, Puerto Rico, and Costa Rica, but remain unavailable
for the central Mexican Caribbean
\citep{barbour_2011,fogg_2013,dahl_2014,aguilarperera_2016,sabidoitza_2016,darling_2011,edwards_2014,chin_2016,deleon_2013,toledohernndez_2014,sandel_2015}.

@@

Length-weight relationships (n = 15) identified in literature were
obtained for the North Atlantic (n = 1), Gulf of Mexico (n = 7,), and
Caribbean (n = 8). Information on sampling methods, sex differentiation,
location, and depth ranges of each study was retrieved when available.
Studies were assumed to include both genders if gender was unspecified.
Locations where allometric studies were performed are shown in Figure
\ref{fig:map} and Table \ref{tab:all_params}.

@@

Studies inconsistently defined a as either the ponderal index from Eq.
\ref{eq:allometric} or the y-intercept (\(c\)) from Eq.
\ref{eq:log-alo-trans}. Other studies incorrectly reported parameters as
mm-to-g conversions when they were in fact cm-to-g conversions. Here,
all parameters are reported as TL(mm) to TW(gr) conversions. When
required, values from other studies were transformed for consistency.

Since uncertainty around estimated relationships was not reported in
some of the reviewed studies, it was not possible to test for
statistical differences between relationships. Instead, the 16
length-weight relationships were used to calculate predicted weight from
length observations of the organisms sampled from the central Mexican
Caribbean. Predicted weights were divided by the observed weights to
obtain a ratio. Difference in mean weight ratios across studies were
tested with a one-way Analysis of Variance (ANOVA). All analyses were
performed in R version 3.4.4 \citep{rcore_2018}. Raw data and code used
in this work are available at dryad.org.

@@

From this study in the central Mexican Caribbean and the 11
peer-reviewed studies that reported length-weight parameters for
\emph{P. volitans} 15 parameters were identified (Table
\ref{tab:all_params}, Fig \ref{fig:all_allo}). Two studies
\citep{aguilarperera_2016,fogg_2013} reported gender-level and pooled
parameters, while the rest presented pooled results. Reviewed studies
presented information for organisms obtained at depths between 0.5 and
57 m. Three studies explicitly stated that their organisms were sampled
with pole spears \citep{aguilarperera_2016,chin_2016,dahl_2014}, and
five studies mentioned that some of their organisms were obtained with
pole spears (or other type of harpoon) but also hand-held nets or fish
traps
\citep{sandel_2015,barbour_2011,fogg_2013,edwards_2014,sabidoitza_2016,toledohernndez_2014},
and two studies did not specify how organisms were sampled
\citep{deleon_2013,darling_2011}. \citet{fogg_2013} use spine-less
weight in the length-weight relationship estimation.

Parameters from models fit to males or females exclusively tend to have
a higher steepness (\emph{i.e.} higher allometric parameter), with mean
\(\pm\) standard deviation values of \(b = 3.27 \pm 0.06\) and
\(b = 3.31 \pm 0.23\) for males and females respectively, compared to
parameters from models for pooled genders with a mean \(\pm\) standard
deviation value of \(b = 3.13 \pm 0.22\). In the case of the ponderal
index (\(a\)) and its \(log_{10}\) transformation (\(c\)), values were
higher for parameters for pooled genders. Figure \ref{fig:all_allo}
shows the length-weight relationships with parameters from all studies.

There were significant differences in predicted-to-observed weight
ratios estimated for each pair of parameters (F(df = 15; 1728) = 38.26;
p \textless{} 0.001). From all allometric parameters reviewed, those of
\citet{edwards_2014} provided the lowest weight estimates, with an
predicted-to-observed weight ratio of 0.98 \(\pm\) 0.23 (mean \(\pm\)
SD). On the other hand,\citet{barbour_2011} yielded the highest weight
estimates, with a mean (\(\pm\) SD) predicted-to-observed weight ratio
of 1.76 \(\pm\) 0.50. Predicted-to-observed weight ratios and groups
identified by Tukey's HSD (\(\alpha = 0.05\)) are presented in Figure
\ref{fig:bio_ratio}.

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{\label{fig:l-w-carib}Length-weight relationship for 109
lionfish sampled in the central Mexican Caribbean. Points indicate
samples, solid line indicates curve of best fit, marginal plots
represent the density distribution of each variable.}
\end{figure}

@@

\caption{\label{tab:unnamed-chunk-5}\label{tab:all_params}Summary of 16 allometric growth parameters available for lionfish in the invaded range from peer-reviewed literature and this study. All parameters have been adjusted to convert from millimeters to grams. n = Sample size, Sex specifies whether data was presented for Females (F), Males (M), or both genders combined (B), a = scaling parameter for Eq. 1 (presented in $\times 10^{-5}$), c = y-intercept for Eq. 3, b = exponent or slope for Eq. 1 or Eq. 3, respectively. The $R^2$ column indicates reported model fit.}
\centering
\begin{tabular}[t]{llllrrll}
\toprule
Region & Sex & n & a & b & c & R\textasciicircum{}2 & Reference\\
\midrule
Caribbean & B & 458 & 3.6 & 2.81 & -4.44 & - & Sandel et al., 2015\\
Caribbean & B & 419 & 2.8 & 2.85 & -4.56 & 0.8715 & Chin et al., 2016\\
Caribbean & B & 1450 & 2.3 & 2.89 & -4.64 & 0.96 & de Leon et al., 2013\\
Caribbean & B & 1887 & 0.3 & 3.24 & -5.52 & 0.97 & Edwards et al., 2014\\
Caribbean & B & - & 0.25 & 3.29 & -5.60 & - & Darling et al., 2011\\
\addlinespace
Caribbean & B & 2143 & 0.52 & 3.18 & -5.28 & 0.9907 & Sabido-Itza et al., 2016\\
Caribbean & B & 227 & 0.8 & 3.11 & -5.10 & 0.958 & Toledo-Hernández et al., 2014\\
Caribbean & B & 109 & 0.32 & 3.23 & -5.49 & 0.9766 & This study\\
GoM & B & 934 & 0.21 & 3.34 & -5.68 & 0.98 & Dahl \& Patterson, 2014\\
GoM & B & 472 & 0.29 & 3.30 & -5.54 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
\addlinespace
GoM & F & 67 & 0.12 & 3.47 & -5.93 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
GoM & M & 59 & 0.42 & 3.23 & -5.38 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
GoM & B & 582 & 0.14 & 3.43 & -5.86 & 0.99 & Fogg et al., 2013\\
GoM & M & 119 & 0.27 & 3.31 & -5.57 & 0.97 & Fogg et al., 2013\\
GoM & F & 115 & 0.68 & 3.14 & -5.17 & 0.94 & Fogg et al., 2013\\
NorthAtlantic & B & 774 & 2.9 & 2.89 & -4.54 & - & Barbour et al.,2011\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
\caption{\label{fig:all_allo}Length-weight relationships (n = 16) for 11
studies and this study. Colors indicate studies from which the
parameters were extracted. Solid lines indicate that the fit was
performed for males and females pooled together. Dotted lines indicate
that the regression was performed on females, and dashed lines indicate
it was performed for males.}
\end{figure}

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
\caption{\label{fig:bio_ratio}Violin plot showing the distribution of
predicted to observed weight ratios for 16 pairs of allometric
parameters. Red and blue circles indicate median and mean values,
respectively. Like letters indicate values that do not differ
significantly (Tukey's HSD; p \textless{} 0.05).}
\end{figure}

@@

There are evident differences in weight-at-length between organisms from
the Caribbean and Gulf of Mexico / North-Western Atlantic. Weight
estimates with parameters from the Gulf of Mexico and North-Western
Atlantic tend to be higher than those from the Caribbean. Similar
regional variation has been reported for age-at-length relationships of
this species across the invaded region \citep{fogg_2015}, or when
comparing populations from the invasion and native ranges
\citep{pusack_2016}. These may be driven by genetic differences or
organisms being exposed to distinct environmental conditions. For
example, work on mitochondrial DNA has shown two distinct population
groups, identified as the ``Caribbean group'' and ``Northern Group''
\citep{betancurr_201}. Alternatively, \citet{fogg_2015} suggest that
differences observed in age-at-length ``may bemore related to climate
rather than other biological and ecological factors''. Differences in
weight-at-length could also reflect differences in energy input
(\emph{i.e.} in some regions, lionfish eat more) or differential usage
of this energy (\emph{e.g.} regional differences in predator abundances
lead to different usage of energy), or a combination of both. Future
research should focus on identifying which occurs here.

The results presented in this paper can have major implications for
management. Our results show that, for a given length, using one set of
parameters or the other can result in a threefold increase in estimated
weight. This becomes especially relevant when estimating biomass
available for harvest, predicting effects on local ecosystems, or
evaluating the effectiveness of removal programs. Using site-specific
values provides a more accurate estimate of fish biomass. Future
research should try to use, to the extent possible, parameters
calculated for their region, or use different parameters to provide
upper and lower bounds in their results. At the same time, this
highlights the need for more basic research that furthers our
understanding of lionfish biology. To better manage the invasion, we
must perform research that can describe biologically important
information of lionfish throughout its invasion range
\citep{johnson_2016}.

>>>>

\begin{abstract}
Lionfish (\emph{Pterois volitans/miles}) are an invasive species in the
Western Atlantic and the Caribbean. In order to better manage the
invasion, we must be able to accurately estimate their total biomass.
This work compares length-weight relationships of the invasive lionfish
through the invasion range. A review of 17 length-weight relationships
reported in 12 peer-reviewed studies shows that lionfish exhibit spatial
variation in weight-at-length. The reviewed parameters indicate that,
for the same length, lionfish in the Caribbean have lower body mass than
in the Atlantic or Gulf of Mexico. This highlights the importance of
using site-specific parameters to estimate biomass from length
observations. This study also reports a new pair of length-weight
parameters (\(a = 3.2056 imes 10^{-6}; b = 3.235\)) for organisms
sampled in the central Mexican Caribbean region. These findings can have
major implications in management, especially when estimating biomass
available for harvest, predicting effects on local ecosystems, or
evaluating the effectiveness of removal programs.
\end{abstract}

@@

Empirical modeling efforts examining the feasibility of establishing
fisheries for lionfish involve modeling changes in biomass in response
to changes in mortality (\emph{i.e.} culling
\citep{cote_2014,barbour_2011,andradibrown_2017}). In this case, biomass
represents the sum of all individual fish weights. The individual weight
(\(W\)) of an organism can be estimated from its length using the
allometric growth equation, \(W = aL^b\), where \(L\) is an organism's
length and \(a\) and \(b\) are constants. This length-weight
relationship for any given species is an essential component of
biomass-based modeling, and can vary across regions as a response to
biotic (\emph{e.g.} local food availability) and abiotic (\emph{e.g.}
water temperature) conditions \citep{johnson_2016}. Site-specific
parameters are therefore necessary in order to obtain accurate biomass
estimates in a given region, especially when research involves
identifying the total biomass available for harvest by fishers
\citep{barbour_2011,morris_2011,johnston_2015,chin_2016}. Equation
parameters for lionfish exist for North Carolina, Northern and Southern
Gulf of Mexico, the Southern Mexican Caribbean, Bahamas, Little Cayman,
Jamaica, Bonaire, Puerto Rico, and Costa Rica, but remain unavailable
for the central Mexican Caribbean
\citep{barbour_2011,fogg_2013,dahl_2014,aguilarperera_2016,sabidoitza_2016,sabidoitz_2016,darling_2011,edwards_2014,chin_2016,deleon_2013,toledohernndez_2014,sandel_2015}.

@@

Length-weight relationships (n = 17) identified in literature were
obtained for the North Atlantic (n = 1), Gulf of Mexico (n = 7,), and
Caribbean (n = 10). Information on sampling methods, sex
differentiation, location, and depth ranges of each study was retrieved
when available. Studies were assumed to include both genders if gender
was unspecified. Locations where allometric studies were performed are
shown in Figure \ref{fig:map} and Table \ref{tab:all_params}.

@@

Studies inconsistently defined \(a\) as either the ponderal index from
Eq. \ref{eq:allometric} or the y-intercept (\(c\)) from Eq.
\ref{eq:log-alo-trans}. Other studies incorrectly reported parameters as
mm-to-g conversions when they were in fact cm-to-g conversions. Here,
all parameters are reported as TL(mm) to TW(gr) conversions. When
required, values from other studies were transformed for consistency.

We compare observed TW to TW estimated with the 18 length-weight
relationsips to quantify the extent to which using \emph{ex situ} this
leads to over- or underestimation. Predicted weights were divided by the
observed weights to obtain a ratio. Difference in mean weight ratios
across studies were tested with a one-way Analysis of Variance (ANOVA).
All analyses were performed in R version 3.5.0 \citep{rcore_2018}. Raw
data and code used in this work are available at dryad.org.

@@

From this study in the central Mexican Caribbean and the 12
peer-reviewed studies that reported length-weight parameters for
\emph{P. volitans} 17 parameters were identified (Table
\ref{tab:all_params}, Fig \ref{fig:all_allo}). Two studies
\citep{aguilarperera_2016,fogg_2013} reported gender-level and pooled
parameters, while the rest presented pooled results. Reviewed studies
presented information for organisms obtained at depths between 0.5 and
57 m. Three studies explicitly stated that their organisms were sampled
with pole spears
\citep{aguilarperera_2016,chin_2016,dahl_2014,sabidoitz_2016}, and five
studies mentioned that some of their organisms were obtained with pole
spears (or other type of harpoon) but also hand-held nets or fish traps
\citep{sandel_2015,barbour_2011,fogg_2013,edwards_2014,sabidoitza_2016,sabidoitz_2016,toledohernndez_2014},
and two studies did not specify how organisms were sampled
\citep{deleon_2013,darling_2011}. \citet{fogg_2013} use spine-less
weight in the length-weight relationship estimation.

Parameters from models fit to males or females exclusively tend to have
a higher steepness (\emph{i.e.} higher allometric parameter), with mean
\(\pm\) standard deviation values of \(b = 3.27 \pm 0.06\) and
\(b = 3.31 \pm 0.23\) for males and females respectively, compared to
parameters from models for pooled genders with a mean \(\pm\) standard
deviation value of \(b = 3.14 \pm 0.20\). In the case of the ponderal
index (\(a\)) and its \(log_{10}\) transformation (\(c\)), values were
higher for parameters for pooled genders. Figure \ref{fig:all_allo}
shows the length-weight relationships with parameters from all studies.

There were significant differences in predicted-to-observed weight
ratios estimated for each pair of parameters (F(df = 15; 1728) = 38.26;
p \textless{} 0.001). From all allometric parameters reviewed, those of
\citet{sabidoitz_2016} in Banco Chinchorro, (Caribbean) provided the
lowest weight estimates, with a predicted-to-observed weight ratio of
0.80 \(\pm\) 0.19 (mean \(\pm\) SD). On the other hand,
\citet{barbour_2011} in the Northern Atlantic yielded the highest weight
estimates, with a mean (\(\pm\) SD) predicted-to-observed weight ratio
of 1.76 \(\pm\) 0.50. Predicted-to-observed weight ratios are presented
in Figure \ref{fig:bio_ratio}.

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{\label{fig:l-w-carib}Length-weight relationship for 109
lionfish sampled in the central Mexican Caribbean. Points indicate
samples, dashed black line indicates curve of best fit, marginal plots
represent the density distribution of each variable.}
\end{figure}

@@

\caption{\label{tab:unnamed-chunk-5}\label{tab:all_params}Summary of 18 allometric growth parameters available for lionfish in the invaded range from peer-reviewed literature and this study. All parameters have been adjusted to convert from millimeters to grams. n = Sample size, Sex specifies whether data was presented for Females (F), Males (M), or both genders combined (B), a = scaling parameter for Eq. 1 (presented in $\times 10^{-5}$), c = y-intercept for Eq. 3, b = exponent or slope for Eq. 1 or Eq. 3, respectively. The $R^2$ column indicates reported model fit.}
\centering
\begin{tabular}[t]{llllrrll}
\toprule
Region & Sex & n & a & b & c & \$R\textasciicircum{}2\$ & Reference\\
\midrule
Caribbean & B & 458 & 3.6 & 2.81 & -4.44 & - & Sandel et al., 2015\\
Caribbean & B & 419 & 2.8 & 2.85 & -4.56 & 0.8715 & Chin et al., 2016\\
Caribbean & B & 1450 & 2.3 & 2.89 & -4.64 & 0.96 & de Leon et al., 2013\\
Caribbean & B & 1887 & 0.3 & 3.24 & -5.52 & 0.97 & Edwards et al., 2014\\
Caribbean & B & - & 0.25 & 3.29 & -5.60 & - & Darling et al., 2011\\
\addlinespace
Caribbean & B & 2143 & 0.52 & 3.18 & -5.28 & 0.9907 & Sabido-Itza et al., 2016\\
Caribbean & B & 227 & 0.8 & 3.11 & -5.10 & 0.958 & Toledo-Hernández et al., 2014\\
Caribbean & B & 449 & 0.23 & 3.25 & -5.64 & 0.97 & Sabido-Itza et al., 2016b\\
Caribbean & B & 368 & 0.32 & 3.19 & -5.50 & 0.98 & Sabido-Itza et al., 2016b\\
Caribbean & B & 109 & 0.32 & 3.23 & -5.49 & 0.9766 & This study\\
\addlinespace
GoM & B & 934 & 0.21 & 3.34 & -5.68 & 0.98 & Dahl \& Patterson, 2014\\
GoM & B & 472 & 0.29 & 3.30 & -5.54 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
GoM & F & 67 & 0.12 & 3.47 & -5.93 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
GoM & M & 59 & 0.42 & 3.23 & -5.38 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
GoM & B & 582 & 0.14 & 3.43 & -5.86 & 0.99 & Fogg et al., 2013\\
\addlinespace
GoM & M & 119 & 0.27 & 3.31 & -5.57 & 0.97 & Fogg et al., 2013\\
GoM & F & 115 & 0.68 & 3.14 & -5.17 & 0.94 & Fogg et al., 2013\\
NorthAtlantic & B & 774 & 2.9 & 2.89 & -4.54 & - & Barbour et al.,2011\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
\caption{\label{fig:all_allo}Length-weight relationships (n = 18) for 12
studies and this study. Colors indicate studies from which the
parameters were extracted. Solid lines indicate that the fit was
performed for males and females pooled together. Dotted lines indicate
that the regression was performed on females, and dashed lines indicate
it was performed for males. The dashed black line represents the
relationship estimated in this study.}
\end{figure}

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
\caption{\label{fig:bio_ratio}Violin plot showing the distribution of
predicted to observed weight ratios for 18 pairs of allometric
parameters. Red and blue circles indicate median and mean values,
respectively. Like letters indicate values that do not differ
significantly (Tukey's HSD; p \textless{} 0.05).}
\end{figure}

\clearpage

@@

There are evident differences in weight-at-length between organisms from
the Caribbean and Gulf of Mexico / North-Western Atlantic. Weight
estimates with parameters from the Gulf of Mexico and North-Western
Atlantic tend to be higher than those from the Caribbean. Similar
regional variation has been reported for age-at-length relationships of
this species across the invaded region \citep{fogg_2015,edwards_2014},
or when comparing populations from the invasion and native ranges
\citep{pusack_2016}. These may be driven by genetic differences or
organisms being exposed to distinct environmental conditions. For
example, work on mitochondrial DNA has shown two distinct population
groups, identified as the ``Caribbean group'' and ``Northern Group''
\citep{betancurr_2011}. Alternatively, \citet{fogg_2015} suggest that
differences observed in age-at-length ``may bemore related to climate
rather than other biological and ecological factors''. Differences in
weight-at-length could also reflect differences in energy input
(\emph{i.e.} in some regions, lionfish eat more) or differential usage
of this energy (\emph{e.g.} regional differences in predator abundances
lead to different usage of energy), or a combination of both. Future
research should focus on identifying which occurs here.

The results presented in this paper can have major implications for
management. Our results show that, for a given length, using one set of
parameters or the other can result in a threefold increase in estimated
weight. Using site-specific values provides a more accurate estimate of
fish biomass. This becomes especially relevant when estimating biomass
available for harvest, predicting effects on local ecosystems, or
evaluating the effectiveness of removal programs. \citet{edwards_2014}
simulated lionfish culling using parameters from North Carolina and
Little Cayman, and identify that the difference in time required for the
population to recover to 90\% of its initial biomass after removals
cease was of up to four years. Future research should try to use, to the
extent possible, parameters calculated for their region, or use
different parameters to provide upper and lower bounds in their results.
At the same time, this highlights the need for more basic research that
furthers our understanding of lionfish biology. To better manage the
invasion, we must perform research that can describe biologically
important information of lionfish throughout its invasion range
\citep{johnson_2016}.
